### 2022-03-05 paper/output/document/thesis_5_discussion.tex
added will's fixes to ch5

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents. My work on automating the categorical imperative 
could serve as one component of a partially or fully artificial ethical reasoner, or an ``ethical AI.''
Specifically, my system is a categorical imperative library that takes as input the logical representation 
of a maxim and returns its moral status (if it is obligatory, prohibited, or permissible).
As it stands, my project can evaluate the moral status of maxims represented in my logic and potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{figure}
\centering
\includegraphics[scale=0.5]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent, which begins with a moral dilemma, 
passes it through an input parser, applies the automated Kantian ethics test, and then processes the 
output using an output parser. I contribute the automated Kantian ethics component.} \label{fig:AIengine}
\end{figure}
%
\begin{isamarkuptext}%
In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work would need to develop such input and output parsers.

@@

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent.
An automated version of the categorical imperative could become part of an ethics engine 
for an AI agent, with additional work to parse the input and the output. Effectively, the kind 
of automated ethics I implement could be a library that AI developers use to give AI agents the capacity for 
sophisticated ethical reasoning faithful to philosophical literature. This represents an improvement 
over existing AI ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In this section, I share an example of the kind of philosophical insight that computational ethics 
can prompt and analyze the value that this tool can offer to philosophers.\footnote{Some may wonder if my system
could also help ordinary people perform ethical reasoning, not just academic philosophers. I address the role that 
my system should play in ordinary ethical reasoning in Appendix \ref{ordinaryhumans} and conclude
that large parts of ordinary ethical inquiry cannot and should not be automated.}%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$. The insight that the FUL should not apply to badly-formed maxims has 
philosophical value and serves as evidence of the potential of computational ethics.

@@

The idea that we cannot even evaluate, let alone will, badly-formed maxims has implications for debates about 
self-doubt and self-respect. The dancing maxim can be read as an expression of self-doubt. Under this 
reading, the question ``When I am dancing, should I be dancing for the sake of dancing?'' is the agent asking, 
``Am I doing the right thing right now?'' Unlike the drowning example, the agent is not asking about the 
next moment, but is expressing doubt about the moral validity of their behavior at this current moment. 
Self-doubt does not always undermine the will—after all, self-doubt plays an 
important role in moral reasoning and is often the mark of a thoughtful agent. Instead, I argue 
that questions of self-doubt do not actually involve badly-formed maxims, for these are not the maxims 
that the agent is doubting. This example will demonstrate that the tension between self-doubt and self-respect 
arises from a mistaken characterization of questions of self-doubt as questions about badly-formed maxims.
I first explain the tension between self-doubt and self-respect in epistemology, then 
explain the parallel tension in ethics, and finally present a resolution of this tension.

@@

The tension between ESR and ESD arises from a misreading of questions of self-doubt as questions about 
the evaluation of badly-formed maxims. A question of self-doubt cannot refer to a badly-formed maxim and must 
instead refer to a well-formed maxim about the agent's past decision-making. As seen before, cases where 
agents appear to ask themselves about badly-formed maxims are mistaken about the maxim in question, because 
such a question could never yield a useful answer for a practical reasoner.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{An Argument For Computational Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The insight above is an example of the kind of philosophical progress that can be 
made using computational tools and serves as evidence for the power of computational ethics. The 
idea that the FUL can only hold for well-formed maxims would have been
incredibly difficult to discover without a computer. I discovered it while formulating the FUL because 
Isabelle's proof-finding tools look for edge cases like badly-formed maxims. Badly-formed maxims are 
interesting because they are the kind of thing that is usually ignored in ordinary philosophical inquiry. 
Philosophers usually assume that we are not discussing badly-formed maxims because, as argued above, 
they are not the kind of thing that is immediately relevant to ethics. Computational tools like Isabelle
require that assumptions like the exclusion of well-formed maxims are made precise, and thus force 
philosophers to understand their arguments in a new way. This example demonstrates the contributions
that computational ethics makes: it can quickly check edge cases and it requires that all assumptions be
made explicit and precise.

@@

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our wildest imaginations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it can 
become as much a tool for philosophy as a calculator is for for arithmetic.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Theoretical Objections to Automating Kantian Ethics \label{amapossible}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Many philosophers cringe at the idea that a computer could perform ethical reasoning or that the 
categorical imperative could provide an algorithm for moral judgement. For example, Rawls asserts, 
``it is a serious misconception to think of the CI-procedure as an algorithm intended to yield, 
more or less mechanically, a correct judgment. There is no such algorithm, and Kant knows this'' \citep[166]{rawlslectures}. 
Ebels-Duggan also claims, ``no one supposes that the Categorical Imperative provides a mechanical 
algorithm that delivers all by itself a complete account of what we ought to do in any given 
situation'' \citep[174]{ebelsduggan}. However unmechanical ethical reasoning
may seem, these claims are not obvious and require further justification. Philosophers who believe 
that mental activity completely determines moral reasoning must explain why computers can, in theory, 
simulate certain mental processes like arithmetic and language, but cannot perform ethical reasoning. 
Without a soul or God-based account of ethical reasoning, it is not obvious that it is theoretically 
impossible to automate ethical reasoning. After all, computers may eventually learn to simulate human 
mental activity entirely, as shown by progress in brain simulation \citep{brainsimulation}.

@@

%:%199=152%:%
%:%200=153%:%
%:%201=154%:%
%:%210=156%:%

@@

%:%395=326%:%
%:%407=328%:%
%:%408=329%:%
%:%409=330%:%
%:%410=331%:%
%:%411=332%:%
%:%412=333%:%
%:%413=334%:%
%:%414=335%:%
%:%415=336%:%
%:%416=337%:%
%:%417=338%:%
%:%418=339%:%
%:%419=340%:%
%:%420=341%:%
%:%421=342%:%

@@

%:%479=394%:%
%:%491=397%:%
%:%492=398%:%
%:%493=399%:%
%:%494=400%:%
%:%495=401%:%
%:%496=402%:%
%:%497=403%:%
%:%498=404%:%
%:%499=405%:%
%:%500=406%:%
%:%501=407%:%
%:%502=408%:%
%:%503=409%:%
%:%504=410%:%
%:%505=411%:%
%:%506=412%:%
%:%507=413%:%
%:%508=414%:%
%:%509=415%:%
%:%510=416%:%
%:%511=417%:%
%:%512=418%:%
%:%513=419%:%
%:%514=420%:%
%:%515=421%:%
%:%516=422%:%
%:%517=423%:%
%:%518=424%:%
%:%519=425%:%
%:%520=426%:%
%:%521=427%:%
%:%522=428%:%
%:%523=429%:%
%:%524=430%:%
%:%525=431%:%
%:%526=432%:%
%:%527=433%:%
%:%528=434%:%
%:%529=435%:%
%:%530=436%:%
%:%531=437%:%
%:%532=438%:%
%:%533=439%:%
%:%534=440%:%
%:%535=441%:%
%:%536=442%:%
%:%537=443%:%
%:%538=444%:%
%:%539=445%:%
%:%540=446%:%
%:%541=447%:%
%:%542=448%:%
%:%543=449%:%
%:%544=450%:%
%:%545=451%:%
%:%546=452%:%
%:%547=453%:%
%:%548=454%:%
%:%549=455%:%
%:%550=456%:%
%:%551=457%:%
%:%552=458%:%
%:%553=459%:%
%:%554=460%:%
%:%555=461%:%
%:%556=462%:%
%:%557=463%:%
%:%558=464%:%
%:%559=465%:%
%:%560=466%:%
%:%561=467%:%
%:%562=468%:%
%:%563=469%:%
%:%564=470%:%
%:%565=471%:%
%:%566=472%:%
%:%567=473%:%
%:%568=474%:%
%:%569=475%:%
%:%570=476%:%
%:%571=477%:%
%:%572=478%:%
%:%573=479%:%
%:%574=480%:%
%:%575=481%:%
%:%576=482%:%
%:%577=483%:%
%:%578=484%:%
%:%579=485%:%
%:%580=486%:%
%:%581=487%:%
%:%582=488%:%
%:%583=489%:%
%:%584=490%:%
%:%585=491%:%
%:%586=492%:%
%:%587=493%:%
%:%588=494%:%
%:%589=495%:%
%:%590=496%:%
%:%591=497%:%
%:%592=498%:%
%:%593=499%:%
%:%594=500%:%
%:%595=501%:%
%:%596=502%:%
%:%597=503%:%
%:%598=504%:%
%:%601=506%:%
%:%602=507%:%
%:%603=508%:%
%:%604=509%:%
%:%605=510%:%
%:%606=511%:%
%:%609=512%:%
%:%610=513%:%
%:%611=514%:%
%:%612=515%:%
%:%613=516%:%
%:%614=517%:%
%:%615=518%:%
%:%616=519%:%
%:%617=520%:%
%:%618=521%:%
%:%619=522%:%
%:%620=523%:%
%:%621=524%:%
%:%622=525%:%
%:%623=526%:%
%:%632=530%:%
%:%644=532%:%
%:%645=533%:%
%:%646=534%:%
%:%647=535%:%
%:%648=536%:%
%:%649=537%:%
%:%650=538%:%
%:%651=539%:%
%:%652=540%:%
%:%653=541%:%
%:%654=542%:%
%:%655=543%:%
%:%656=544%:%
%:%657=545%:%
%:%658=546%:%
%:%659=547%:%
%:%660=548%:%
%:%661=549%:%
%:%662=550%:%
%:%663=551%:%
%:%664=552%:%
%:%665=553%:%
%:%666=554%:%
%:%667=555%:%
%:%668=556%:%
%:%669=557%:%
%:%670=558%:%
%:%671=559%:%
%:%672=560%:%
%:%673=561%:%
%:%674=562%:%
%:%675=563%:%
%:%676=564%:%
%:%677=565%:%
%:%678=566%:%
%:%679=567%:%
%:%680=568%:%
%:%681=569%:%
%:%682=570%:%
%:%683=571%:%
%:%684=572%:%
%:%685=573%:%
%:%686=574%:%
%:%687=575%:%
%:%688=576%:%
%:%689=577%:%
%:%690=578%:%
%:%691=579%:%
%:%692=580%:%
%:%693=581%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{thesis{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}discussion}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Discussion\label{discussion}%
}
\isamarkuptrue%
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Chapter \ref{applications}, I demonstrated that my system is capable of performing sophisticated,
nuanced ethical reasoning. In this section, I outline the additional components necessary for my system
to guide AI agents. My work on automating the categorical imperative 
could serve as one component of a partially or fully artificial ethical reasoner, or an ``ethical AI.''
Specifically, my system is a categorical imperative library that takes as input the logical representation 
of a maxim and returns its moral status (if it is obligatory, prohibited, or permissible).
As it stands, my project can evaluate the moral status of maxims represented in my logic and potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, my system could be combined with an input parser to translate moral dilemmas as represented 
to the AI agent into maxims in my logic. An output parser could translate my system's output
into a prescription for action that the AI agent could act on. Figure \ref{fig:AIengine} depicts the 
workflow of this example ethics engine.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{figure}
\centering
\includegraphics[scale=0.5]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent, which begins with a moral dilemma, 
passes it through an input parser, applies the automated categorical imperative test, and then processes the 
output using an output parser. I contribute the automated categorical imperative component.} \label{fig:AIengine}
\end{figure}
%
\begin{isamarkuptext}%
In this workflow, an AI agent is faced with a moral dilemma in some internal representation. The input
parser translates this internal representation into an appropriate logical representation, i.e. 
a circumstance, act, goal tuple. The output parser translates the output of the categorical imperative
library (the moral status of the maxim as obligatory, prohibited, or permissible) to a prescription for
action. In order for my system to be used in an AI agent using the workflow
above, future work would need to develop such input and output parsers.

@@

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent.
An automated version of the categorical imperative could become part of an ethics engine 
for an AI agent, with additional work to parse the input and the output. Effectively, the kind 
of automated ethics I implement could be a library that AI developers use to give AI agents the capacity for 
sophisticated ethical reasoning faithful to philosophical literature. This represents an improvement 
over existing AI ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help human beings make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In this section, I share an example of the kind of philosophical insight that computational ethics 
can prompt and analyze the value that this tool can offer to philosophers.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$. The insight that the FUL should not apply to badly-formed maxims has 
philosophical value and serves as evidence of the potential of computational ethics.

@@

In the previous section, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I demonstrate the implications of this concept for a 
philosophical debate to demonstrate that computational ethics can result in insights with real 
philosophical weight. I use badly formed maxims to resolve the tension between self-doubt and self-respect. 
This debate originated in epistemology, which values doubting your beliefs but also rationally requires
that you believe that you are not mistaken (else you should update your beliefs). 
Self-doubt and self-respect also compete in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves, and it becomes clear that both concepts can
co-exist, as we intuitively desire. I sketch the details of this argument in the rest of this section.

Another reading of the dancing maxim understands it as an expression of self-doubt. Under this 
reading, the question ``When I am dancing, should I be dancing for the sake of dancing?'' is the agent asking, 
``Am I doing the right thing right now?'' Unlike the drowning example, the agent is not asking about the 
next moment, but is expressing doubt about the moral validity of their behavior at this current moment. 
Surely self-doubt does not always undermine the will—after all, self-doubt plays an 
important role in moral reasoning and is often the mark of a thoughtful agent. I argue 
that questions of self-doubt do not actually involve badly-formed maxims, for these are not the maxims 
that the agent is doubting. I first frame this debate by introducing the tension between self-doubt and 
self-respect in epistemology, then explain the parallel tension in ethics, and finally present a 
resolution of this tension.

@@

The tension between ESR and ESD arises from a misreading of questions of self-doubt as questions about 
the evaluation of badly-formed maxims. A question of self-doubt cannot refer to a badly-formed maxim and must 
instead refer to a well-formed maxim about the agent's past decision-making. As seen before, cases where 
agents appear to ask themselves about badly-formed maxims are mistaken about the maxim in question, because 
such a question could never yield a useful answer for a practical reasoner.

By recognizing that the naive version of ethical self-doubt is a badly formed maxim, I realized 
that there is something wrong with its formulation and modified it to resolve the above debate. 
This demonstrates a larger meta-pattern in Kantian reasoning: many debates in Kantian philosophical
literature revolve around incorrectly formulated maxims, so the insight about badly
formed maxims is significant. Much of the work of a Kantian ethicist is formulating an
appropriate maxim, and badly-formed maxims define one boundary condition for this task.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{An Argument For Computational Ethics%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
The insight above is an example of the kind of philosophical progress that can be 
made using computational tools and serves as evidence for the power of computational ethics. The 
idea that the FUL can only hold for well-formed maxims would have been
incredibly difficult to discover without a computer. I discovered it while formulating the FUL because 
Isabelle's proof-finding tools look for edge cases like badly-formed maxims. Badly-formed maxims are 
interesting because they are the kind of thing that is usually ignored in ordinary philosophical inquiry. 
Philosophers usually assume that we are not discussing badly-formed maxims because, as argued above, 
they are not the kind of thing that is immediately relevant to ethics. Computational tools like Isabelle
require that assumptions like the exclusion of well-formed maxims are made precise, and thus force 
philosophers to understand their arguments in a new way. This example demonstrates the contributions
that computational ethics makes: it can quickly check edge cases and it requires that all assumptions be
made explicit and precise.

@@

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our wildest imaginations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it could
become as commonplace in philosophy departments as reflective equilibrium. Just as computational tools
have supercharged progress in healthcare and drug discovery, computational ethics has the potential for
great philosophical progress.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automating Ordinary Ethical Reasoning \label{ordinarypeople}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Sections \ref{AIethics} and \ref{computationalethics}, I outline how automated ethics can guide
artificial agents and human philosophers respectively. This raises a natural question: can automated
ethics guide ordinary human beings, not just academic philosophers, as we navigate the world and face ethical
dilemmas? Some may hope (or worry) that automated ethics could render ethical reasoning obsolete. In 
this section, I argue, that while computers should not replace human ethical reasoning entirely,
they can supplement and improve our ethical reasoning. I argue for a kind of human-computer symbiosis
in which computers offer ethical advice, arguments for particular moral judgements, and speed up moral
calculations without subverting human ethical reasoning entirely \citep{licklider}.

Ethics bears weight for everyone, not just for academic philosophers, because it studies the unavoidable question:
how should we live? If computers can make this study more efficient, then it seems that everyone should
engage in computational ethics. The ethical question is the only question that 
we answer merely by living. To turn away from ethics is to take a stance on the question of how to 
live (namely, to live unreflectively) and thus to engage in ethics. Every rational being must decide 
how to navigate the world and ethics answers this question. Given that ethics is vital, it seems that if
comptuational tools can help us derive ethical judgements more efficiently, then we should automate as 
much ethical reasoning as possible.
In the most extreme case, we can unthinkingly follow the commands of an ethical calculator that dictates 
how we should live. Maybe computers can answer the unavoidable question for us.

The argument above places the value of ethics solely in its action-guiding potential, and thus fails to take into account the 
importance of practical reason, which, as I argued in Section \ref{kantianethics}, is the source
of freedom itself. We are committed to ethical reflection because of the kind of beings that we are. 
Recall that Korsgaard argues that, as beings occupying minds with a reflective structure, when faced with 
a choice,``it is as if there were something over and above all of your desires, something that is you, and that chooses which desire 
to act on'' \citep[83]{sources}. This choosing is the operation of practical reason, and this reflection
makes us free. We are free because we must choose which reasons to act on. Every decision that we 
make is an exercise of freedom.

If reflection makes us free, then unthinkingly obeying a computer sacrifices our autonomy. Consider 
an Ethics Oracle that can unfailingly tell you the right thing to do in any 
situation.\footnote{This example is inspired by the Pocket Oracle presented in \citet{bok}.} Someone 
who surrenders themselves to this Oracle unthinkingly follows its prescriptions. 
The reflection involved in the decision to obey each of the Oracle’s prescriptions is limited \citep{bok}. 
This person is not reflecting on the real matters at hand and is 
not making decisions for themselves. They have surrendered their reflective capacity to the Oracle. 
They live a worse life than someone who reflects on their actions; they have less ownership over their 
actions than the reflective person. In a less extreme case, a person may retain control of many of 
their decisions but cede some important or tricky choices to the Ethics Oracle. Because every single 
exercise of practical reason is an exercise of autonomy, this person is still less autonomous than the 
purely reflective person. Even surrendering simple, inconsequential decisions such as which flavor of 
coffee to drink surrenders some piece of our autonomy. Perhaps in trivial cases we can accept that 
tiny sacrifice, but giving over life-changing decisions to the machine sacrifices our 
core freedom. Unreflectively relying on computational ethics surrenders our autonomy to the machine.

One objection to this emphasis on reflection is the impracticality of making ethical calculations from first principles 
every time we are faced with a decision. This is why we follow the advice of moral mentors, like our 
family or influential philosophers. 
Most people do not reason about ethics during everyday decisions; they rely on some combination of 
prior knowledge and external testimony. For example, my mother taught me to respect myself, so I 
follow her advice.

What is the difference between following the guidance of a moral educator and
obeying the Ethics Oracle? The best kind of ethical advice prompts reflection, such as an argument 
made in a philosophy paper. Unthinkingly following someone’s advice results in the same loss of 
autonomy as unthinkingly obeying the Ethics Oracle; people who merely obey orders are less autonomous 
than those who think for themselves. This account of moral advice offers a model for human-computer 
symbiotic computational ethics. The computer should serve as a moral guide by providing arguments, just as my mom 
explained why I should always respect myself. Human-computer symbiotic computational
ethics nurtures autonomy when it not only offers prescriptions for action, but also explanations for 
these prescriptions. Because my theorem-prover-based automated ethical system is explainable, it can 
guide action without sacrificing autonomy. It can make an argument for some action, instead of 
merely giving a verdict. Isabelle can list the facts used to show
a partcular action prohibited, and a human being can reflect on whether or not these principles
indeed prohibit the action in question. The computer serves as a collaborator and a tool, but not as an authority, 
so the human being’s reflective capacity and freedom is preserved.

The above model of human-computer symbiosis demonstrates how computational ethics can augment human
ethical reasoning without replacing it. When deliberating over moral dilemmas, ordinary people can 
turn to computational tools for advice, like an ``Ask an Ethicist'' column. If we appeal to philosophically
faithful computational ethics like that implemented in this thesis, then this advice will synthesize 
decades of philosophical progress and is thus a way to apply philosophers' insights to ordinary life. 
Moreover, just as they do for philosophers, computers can help ordinary people approach ethical questions from a 
different perspective. Even interacting with my system requires the user to consider the action's maxim, 
which includes the circumstances, act, and goal. Making these components of action precise already changes
the user's perspective. Just as computational ethics can serve as a tool for academic philosophers to 
automate away tedium and achieve greater precision, it can do the same for ordinary human beings navigating
the world. Moreover, it also offers another way for the general public to access professional philosophy's
insights, and thus carries potential to improve our everyday reasoning.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Theoretical Objections to Automating Kantian Ethics \label{amapossible}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Many philosophers cringe at the idea that a computer could perform ethical reasoning or that the 
categorical imperative could provide an algorithm for moral judgement. For example, Rawls asserts, 
``it is a serious misconception to think of the CI-procedure as an algorithm intended to yield, 
more or less mechanically, a correct judgment. There is no such algorithm, and Kant knows this'' \citep[166]{rawlslectures}. 
Ebels-Duggan also claims, ``no one supposes that the Categorical Imperative provides a mechanical 
algorithm that delivers all by itself a complete account of what we ought to do in any given 
situation'' \citep[174]{ebelsduggan}. However unmechanical ethical reasoning
may seem, these claims are not obvious and require further justification. Philosophers who believe 
that mental activity completely determines moral reasoning must explain why computers can, in theory, 
simulate certain mental processes like arithmetic and language, but cannot perform ethical reasoning. 
Without a soul or God-based account of ethical reasoning, it is not obvious that it is theoretically 
impossible to automate ethical reasoning. After all, computers may eventually learn to simulate human 
mental activity entirely, as shown by progress in brain simulation \citep{brainsimulation}.

@@

%:%207=153%:%
%:%219=155%:%
%:%220=156%:%
%:%221=157%:%

@@

%:%387=323%:%
%:%388=324%:%
%:%389=325%:%
%:%390=326%:%
%:%391=327%:%
%:%392=328%:%
%:%393=329%:%
%:%394=330%:%
%:%395=331%:%
%:%396=332%:%
%:%397=333%:%
%:%398=334%:%
%:%399=335%:%
%:%400=336%:%
%:%401=337%:%
%:%410=341%:%

@@

%:%471=392%:%
%:%472=393%:%
%:%473=394%:%
%:%474=395%:%
%:%475=396%:%
%:%476=397%:%
%:%477=398%:%
%:%478=399%:%
%:%479=400%:%
%:%480=401%:%
%:%481=402%:%
%:%482=403%:%
%:%483=404%:%
%:%484=405%:%
%:%485=406%:%
%:%486=407%:%
%:%487=408%:%
%:%496=410%:%
%:%508=412%:%
%:%509=413%:%
%:%510=414%:%
%:%511=415%:%
%:%512=416%:%
%:%513=417%:%
%:%514=418%:%
%:%515=419%:%
%:%516=420%:%
%:%517=421%:%
%:%518=422%:%
%:%519=423%:%
%:%520=424%:%
%:%521=425%:%
%:%522=426%:%
%:%523=427%:%
%:%524=428%:%
%:%525=429%:%
%:%526=430%:%
%:%527=431%:%
%:%528=432%:%
%:%529=433%:%
%:%530=434%:%
%:%531=435%:%
%:%532=436%:%
%:%533=437%:%
%:%534=438%:%
%:%535=439%:%
%:%536=440%:%
%:%537=441%:%
%:%538=442%:%
%:%539=443%:%
%:%540=444%:%
%:%541=445%:%
%:%542=446%:%
%:%543=447%:%
%:%544=448%:%
%:%545=449%:%
%:%546=450%:%
%:%547=451%:%
%:%548=452%:%
%:%549=453%:%
%:%550=454%:%
%:%551=455%:%
%:%552=456%:%
%:%553=457%:%
%:%554=458%:%
%:%555=459%:%
%:%556=460%:%
%:%557=461%:%
%:%558=462%:%
%:%559=463%:%
%:%560=464%:%
%:%561=465%:%
%:%562=466%:%
%:%563=467%:%
%:%564=468%:%
%:%565=469%:%
%:%566=470%:%
%:%567=471%:%
%:%568=472%:%
%:%569=473%:%
%:%570=474%:%
%:%571=475%:%
%:%572=476%:%
%:%573=477%:%
%:%574=478%:%
%:%575=479%:%
%:%576=480%:%
%:%577=481%:%
%:%578=482%:%
%:%579=483%:%
%:%580=484%:%
%:%581=485%:%
%:%582=486%:%
%:%583=487%:%
%:%584=488%:%
%:%585=489%:%
%:%586=490%:%
%:%595=493%:%
%:%607=496%:%
%:%608=497%:%
%:%609=498%:%
%:%610=499%:%
%:%611=500%:%
%:%612=501%:%
%:%613=502%:%
%:%614=503%:%
%:%615=504%:%
%:%616=505%:%
%:%617=506%:%
%:%618=507%:%
%:%619=508%:%
%:%620=509%:%
%:%621=510%:%
%:%622=511%:%
%:%623=512%:%
%:%624=513%:%
%:%625=514%:%
%:%626=515%:%
%:%627=516%:%
%:%628=517%:%
%:%629=518%:%
%:%630=519%:%
%:%631=520%:%
%:%632=521%:%
%:%633=522%:%
%:%634=523%:%
%:%635=524%:%
%:%636=525%:%
%:%637=526%:%
%:%638=527%:%
%:%639=528%:%
%:%640=529%:%
%:%641=530%:%
%:%642=531%:%
%:%643=532%:%
%:%644=533%:%
%:%645=534%:%
%:%646=535%:%
%:%647=536%:%
%:%648=537%:%
%:%649=538%:%
%:%650=539%:%
%:%651=540%:%
%:%652=541%:%
%:%653=542%:%
%:%654=543%:%
%:%655=544%:%
%:%656=545%:%
%:%657=546%:%
%:%658=547%:%
%:%659=548%:%
%:%660=549%:%
%:%661=550%:%
%:%662=551%:%
%:%663=552%:%
%:%664=553%:%
%:%665=554%:%
%:%666=555%:%
%:%667=556%:%
%:%668=557%:%
%:%669=558%:%
%:%670=559%:%
%:%671=560%:%
%:%672=561%:%
%:%673=562%:%
%:%674=563%:%
%:%675=564%:%
%:%676=565%:%
%:%677=566%:%
%:%678=567%:%
%:%679=568%:%
%:%680=569%:%
%:%681=570%:%
%:%682=571%:%
%:%683=572%:%
%:%684=573%:%
%:%685=574%:%
%:%686=575%:%
%:%687=576%:%
%:%688=577%:%
%:%689=578%:%
